### 2013-02-12 doc/xps2013/header.tex
partial commit for Luis the sick boy
done for the day
4l

@@

>>>>

\newcommand{\BigO}[1]{$O(#1)$}

\newcommand{\api}{\textls*{\small API}}
\newcommand{\apis}{\textls*{\small API}\hspace{0.75pt}s}
\newcommand{\ssd}{\textls*{\small SSD}}
\newcommand{\ssds}{\textls*{\small SSD}\hspace{0.75pt}s}

@@

\makeatletter
\renewcommand{\paragraph}{%
  \@startsection{paragraph}{4}%
  {\z@}{1.0ex \@plus 1ex \@minus .2ex}{-1em}%
  {\normalfont\normalsize\bfseries}%
}

\renewcommand\section{\@startsection{section}{1}{\z@}{-10\p@ \@plus -4\p@ \@minus -4\p@}{5\p@ \@plus 4\p@ \@minus 4\p@}{\normalfont\Large\bfseries\boldmath\rightskip=\z@ \@plus 8em\pretolerance=10000 }}

\renewcommand\subsection{\@startsection{subsection}{2}{\z@}{-10\p@ \@plus -4\p@ \@minus -4\p@}{5\p@ \@plus 4\p@ \@minus 4\p@}{\normalfont\large\bfseries\boldmath\rightskip=\z@ \@plus 8em\pretolerance=10000 }}

\renewcommand\subsubsection{\@startsection{subsubsection}{3}{\z@}{-10\p@ \@plus -4\p@ \@minus -4\p@}{5\p@ \@plus 4\p@ \@minus 4\p@}{\normalfont\normalsize\bfseries\boldmath\rightskip=\z@ \@plus 8em\pretolerance=10000 }}

\makeatother

### 2013-02-12 doc/xps2013/proposal.tex
4l

Irregular applications generate tasks with work, dependences, and memory
accesses that are highly sensitive to input. Classic examples of irregular
applications include branch and bound optimization, SPICE circuit simulation,
contact algorithms in car crash analysis, and network flow. Important
contemporary examples include processing large graphs in the business,
national security, machine learning, data-driven science, and social network
computing domains. For these emerging applications, fast response -- given the
sheer amount of data -- requires multinode systems. The most broadly available
multinode systems are those built from x86 commodity computing nodes
interconnected via ethernet or infiniband. Our current work enables scalable
performance of irregular applications on these mass market systems,
addressing two key challenges:

@@

\begin{itemize}
\item \textbf What is the right system architecture support for irregular applications, including architecture, communication and storage?
\item \textbf What are the necessary language and compiler support for large-scale, data-intensive irregular applications?
\item \textbf How do we comprehensively evaluate performance of graph analytics systems?
\end{itemize}

>>>>

Irregular applications generate tasks with work, dependences, and memory
accesses that are highly sensitive to input. Classic examples of irregular
applications include branch and bound optimization, SPICE circuit simulation,
contact algorithms in car crash analysis, and network flow. Important
contemporary examples include processing large graphs in the business,
national security, machine learning, data-driven science, and social network
computing domains. For these emerging applications, fast response -- given the
sheer amount of data -- requires multinode systems. The most broadly available
multinode systems are those built from x86 commodity computing nodes
interconnected via ethernet or infiniband. Our focus is to enable scalable
performance of irregular applications on these mass market systems,
addressing two key challenges:

@@

\begin{itemize}
\item \textbf{What is the right system architecture support for irregular applications, including architecture, communication and storage?}
\item \textbf{What are the necessary language and compiler support for large-scale, data-intensive irregular applications?}
\item \textbf{How do we comprehensively evaluate performance of graph analytics systems?}
\end{itemize}

\paragraph{Intellectual Merit:}

Efficiently executing irregular computations on commodity hardware is \emph{hard}.  Much about commodity hardware is seemingly designed for the opposite application characteristics: caches assume applications have locality; networks are built with low injection rates but high bandwidth, assuming large packet sizes; even the runtimes and operating system is built assuming relatively infrequent communication and synchronization compared to computation.  This proposal rests on the seemingly outlandish claim that this commodity hardware / software stack can be convinced, through another layer of software, into efficiently executing irregular applications that lack locality, and communicate frequently with small amounts of data.  A folksy way of saying it is we think we can jam a square peg through a round hole.  Fortunately we have some preliminary evidence we won't get wedged along the way.  This is an intellectually challenging endeavor, and one square in the middle of the NSF XPS program: using and exploiting parallelism from irregular computations to efficiently execute them at scale.

\paragraph{Broader Impact:}

Irregular computations are of increasing economic importance.  Moreover, they are key to several national security applications.  If commodity hardware can be cajoled into executing such applications efficiently it will open the door for new and rich applications.  Graph queries, machine-learning and data-driven science will be enabled by the technology we aim to create from the ideas in this grant proposal.  In addition our work will include the construction of a comprehensive irregular benchmark suite, something to help guide our own and the field's research.
